### 2016-06-18 opsem.tex
Typos.

Small-step semantics of \txnimp are defined in terms of an abstract
machine that generates an execution trace, also called an execution
($\E$). First artifact of the trace is a set ($\A$) of \emph{effects},
where an effect ($\eta$) documents a read (\C{RD(X)}), or a write
(\C{WR(X)}), or a transaction commit (\C{COMMIT}) operation performed
during the execution. Any value associated with the operation (\eg,
value read or value written) is also documented, and can be accessed
via $\rval$. Every effect has a unique identifier accessible via
$\id$, and a transction idenfier accessible via $\txn$.  The latter
identifies the transaction that generated the effect. In every step of
the evaluation, abstract machine reduces a \txnimp term by executing a
read, write or commit operation, generating an effect, and extending
the trace. Since effects include transaction identifiers, semantics
distinguishes between terms ($t$) of different transactions. For
example, $\txnbox{t}_i$ denotes a term $t$ inside a transaction $T_i$.
Evaluation contexts are also appropriately marked. For example,
$\ectx_i$ denotes evaluation context for a term inside $T_i$. The
other component of the execution trace is the visibility relation
($\visZ$) between the effects. The intent and mechanics of $\visZ$
become clear as we describe the evaluation rules in the following.

Fundamental to our operational semantics is the notation of a trace
invariant ($\I$), which is a function from traces ($\E$) to
propositions ($\texttt{Prop}$) that declare well-formedness
constraints over traces. Abstract machine takes a step only if the
resulting trace satisfies the well-formedness constraints imposed by
$\I$. This behaviour is captured by the auxiliary reduction rule
\rulelabel{E-Aux} that factors out the trace extension aspect of the
evaluation by abstracting away operation-specific behaviour as a
function ($\mathcal{F}$) that generates appropriate effect.
\rulelabel{E-Aux} uses $\mathcal{F}$ to generate a new effect and
extend the trace ($\E = (\A,\visZ)$) \emph{only if} the
well-formedness constraints imposed by $\I$ on $\E$ (i.e., $\I(\E')$)
are satisfied. Otherwise, it gets stuck. In an execution that runs to
completion, every small-step preserves well-formedness of trace, thus
ensuring the invariance of $\I$. Note that the operational semantics
makes no assumptions about $\I$ other than its type. As such it is
parametric over $\I$, and can be instantiated with any
trace-parametric proposition that expresses constraints over the given
trace. While the machine takes a step only if the constraints are
satisfied, it neither defines nor explicitly assumes an oracle to
check the satisfaction; there is no motivation for doing so in the
semantics.

>>>>

Small-step semantics of \txnimp are defined in terms of an abstract
machine that generates an execution trace ($\E$). First artifact of the trace is a set ($\A$) of \emph{effects},
where an effect ($\eta$) documents a read (\C{RD(X)}), or a write
(\C{WR(X)}), or a transaction commit (\C{COMMIT}) operation performed
during the execution. Any value associated with the operation (\eg,
value read or value written) is also documented, and can be accessed
via $\rval$. Every effect has a unique identifier accessible via
$\id$, and a transction identifier accessible via $\txn$.  The latter
identifies the transaction that generated the effect. In every step of
the evaluation, abstract machine reduces a \txnimp term by executing a
read, write or commit operation, generating an effect, and extending
the trace. Since effects include transaction identifiers, semantics
distinguishes between terms ($t$) of different transactions. For
example, $\txnbox{t}_i$ denotes a term $t$ inside a transaction $T_i$.
Evaluation contexts are also appropriately marked. For example,
$\ectx_i$ denotes evaluation context for a term inside $T_i$. The
other component of the execution trace is the visibility relation
($\visZ$) between the effects. The intent and mechanics of $\visZ$
become clear as we describe the evaluation rules in the following.

Fundamental to our operational semantics is the notion of a trace
invariant ($\I$), which is a function from traces ($\E$) to
propositions ($\texttt{Prop}$) that declare well-formedness
constraints over traces. Abstract machine takes a step only if the
resulting trace satisfies the well-formedness constraints imposed by
$\I$. This behaviour is captured by the auxiliary reduction rule
\rulelabel{E-Aux} that factors out the trace extension aspect of the
evaluation by abstracting away operation-specific behaviour as a
function ($\mathcal{F}$) that generates appropriate effect.
\rulelabel{E-Aux} uses $\mathcal{F}$ to generate a new effect and
extend the trace ($\E = (\A,\visZ)$) \emph{only if} the
well-formedness constraints imposed by $\I$ on $\E$ (i.e., $\I(\E')$)
are satisfied. Otherwise, it gets stuck. In an execution that runs to
completion, every small-step preserves well-formedness of trace, thus
ensuring the invariance of $\I$. Note that the operational semantics
makes no assumptions about $\I$ other than its type. As such it is
parametric over $\I$, and can be instantiated with any
trace-parametric proposition that expresses constraints over the given
trace. While the machine takes a step only if the constraints are
satisfied, it neither defines nor explicitly assumes an oracle to
check the satisfaction; there is no need for doing so in the
semantics.
